Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-16 meta_RL_analyses.Rmd
integration stuff

```{r}
summary_plot(reward_d %>%
               filter(run_type %in% c("basic_slower_integration", "basic_random_holdout_longer_slower_integration"),
                      epoch < 500),
             "epoch", "reward", "is_new_game",
             continuous_x=T) +
  facet_wrap(. ~ named_run_type) +
  geom_hline(yintercept=mean(optimal_baselines$expected_reward),
             alpha=0.5) +
  labs(x = "Epoch after introducing new tasks") +
  theme(plot.margin=margin(t=0.24, b=1.7, unit="cm"))

>>>>

```{r}
summary_plot(reward_d %>%
               filter(run_type %in% c("basic_slower_integration", "basic_random_holdout_longer_slower_integration"),
                      epoch < 500),
             "epoch", "reward", "is_new_game",
             continuous_x=T) +
  facet_wrap(. ~ named_run_type) +
  geom_hline(yintercept=mean(optimal_baselines$expected_reward),
             alpha=0.5) +
  labs(x = "Epoch after introducing new tasks")

### 2019-05-16 writing/embedded_meta_learning.tex
integration stuff

\[\hat{z}_{new} = f_{z_{\text{function}}}(z_{old}) \qquad \text{where } f_{z_{\text{function}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]
For example, if the $z_{old}$ is the system's embedding for trying to win at go, and $z_{meta}$ is the meta-task of switching to trying to lose a game, then $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate how well the system actually loses with the $\hat{z}_{new}$ strategy. \par

@@

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\section{Learning multivariate polynomials}
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/poly/basic_results.png}
\caption{The system solves the meta-learning problem and successfully generalizes to held out tasks. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). Error-bars throughout are bootstrap 95\%-confidence intervals.}
\label{poly_basic_results}
\end{figure}

@@

\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The solid line indicates optimal performance; the dashed line indicates chance (untrained). Error-bars are bootstrap 95\%-CIs.} 

@@

\paragraph{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of their application or receiving a natural-language instruction. \par 

@@

\section{A stochastic learning setting: simple card games}
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}
\caption{Generalization to held-out tasks.}
\label{cards_basic_results}
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning_integrating_over_time.png}
\caption{Rapid integration once held-out tasks are introduced.}
\label{cards_basic_integration_results}
\caption{Basic meta learning in the cards domain. (a) The system successfully generalizes to held out tasks, both when trained on a random sample of half the tasks, or when a targeted subset is held out. The gray dashed line indicates chance performance, while the solid lines are optimal performance. The colored dashed lines show how well the system could perform by playing the strategy from the most correlated trained task. The fact that it generally exceeds this difficult baseline shows a deeper form of generalization than just memorizing a few strategies and picking the closest. (b) The system can rapidly adapt to new tasks. (This figure shows the effect of training all parameters while replaying old tasks, unlike the continual learning strategy above.) Error-bars are bootstrap 95\%-CIs.}

@@

\paragraph{Basic meta learning:} First, we show that the system is able to achieve the basic goal of playing a held-out game from examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. Note that the sample complexity in terms of number of trained tasks is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. Finally, the system can integrate new tasks rapidly (fig. \ref{cards_basic_integration_results}).\par

@@

@@

\section{Numerical results}
In this section we provide the mean values and bootstrap confidence intervals corresponding to the major figures in the paper, as well as the baseline results in those figures. Tables were generated with stargazer \citep{Hlavac2018}.

>>>>

\[\hat{z}_{new} = f_{z_{\text{function}}}(z_{old}) \qquad \text{where } f_{z_{\text{meta}}} \text{ is parameterized by } \mathcal{H}\left(z_{\text{meta}}\right)\]
For example, if the $z_{old}$ is the system's embedding for trying to win at go, and the meta-dataset consists of examples of a winning strategy and a losing strategy for different games, then $\hat{z}_{new}$ could be interpreted as the system's guess at a losing strategy for go. We could then evaluate how well the system actually loses with the $\hat{z}_{new}$ strategy. \par

@@

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\section{Learning multivariate polynomials} \label{sec_poly}
\begin{figure}
\centering
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/basic_results.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{poly_basic_results}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{cards_basic_results}
\end{subfigure}%
\caption{The EML system succeeds at the basic meta-learning tasks, which are a necessary prerequisite for meta-mappings. (\subref{poly_basic_results}) Basic meta-learning in the polynomials domain, section \ref{sec_poly}. The system successfully generalizes to held out polynomials. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). (\subref{cards_basic_results}) Basic meta-learning in the cards domain, section \ref{sec_cards}. The system successfully generalizes to held out games, both when trained on a random sample of half the tasks, or when a targeted subset is held out. The gray dashed line indicates chance performance, while the solid lines are optimal performance. The colored dashed lines show how well the system could perform by playing the strategy from the most correlated trained task. The fact that it generally exceeds this difficult baseline shows a deeper form of generalization than just memorizing a few strategies and picking the closest. Error-bars throughout are bootstrap 95\%-confidence intervals, numerical values for plots can be found in appendix \ref{app_numerical_results}.}
\end{figure}

@@

\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). This shows the loss produced when evaluating the mapped embedding on the target task. For example, if the initial polynomial is $p(x) = x + 1$, and the meta-task is ``square,'' the loss would be evaluated by taking the embedding produced by applying this meta-mapping to the embedding of $f(x)$ and evaluating how the mapped embedding regresses $p(x)^2 = x^2 + 2x + 1$. The solid line indicates optimal performance; the dashed line indicates chance (untrained). Error-bars are bootstrap 95\%-CIs.} 

@@

\paragraph{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. We train the system to perform a variety of mappings, for example switch the first two inputs of the polynomial, add 3 to the polynomial, or square the polynomial. We then test its ability to generalize to held-out mappings, for example a permutation of the polynomial inputs which it has not encountered before, or an additive shift in the polynomial which it has not been trained on. The system is both able to apply its learned meta-mappings to held-out tasks, and to apply meta-mappings it has not been trained on, simply by seeing examples of their application or receiving a natural-language instruction. However, it generalizes better to new meta-mappings from examples than language, which is not completely surprising, since the examples provide a richer dataset to learn from. \par 

@@

\section{A stochastic learning setting: simple card games}\label{sec_cards}

@@

\paragraph{Basic meta learning:} First, we show that the system is able to achieve the basic goal of playing a held-out game from examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. Note that the sample complexity in terms of number of trained tasks is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. \par

@@

The results reported in the figures in this paper are averages across multiple runs, with different trained and held-out tasks (in the polynomial case) and different network initializations (in all cases), to ensure the robustness of the findings. \par

@@

\begin{figure}[H]
\centering
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/basic_results_over_time.png}
\caption{The polynomials domain, section \ref{sec_poly}.}
\label{supp_poly_integration_results}
\end{subfigure}%
\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning_integrating_over_time.png}
\caption{The cards domain, section \ref{sec_cards}.}
\label{supp_cards_integration_results}
\end{subfigure}%
\caption{Integrating new tasks into the system by training all parameters results in some initial interference with prior tasks (even with replay), suggesting that an approach like the continual learning-approach may be useful.}
\label{supp_integration_results}
\end{figure}

\section{Numerical results} \label{app_numerical_results}
In this section we provide the mean values and bootstrap confidence intervals corresponding to the major figures in the paper, as well as the baseline results in those figures. Tables were generated with stargazer \citep{Hlavac2018}. \par

\subsection{Polynomials}
\begin{table}[H]
\scriptsize
\centering
\begin{tabular}{@{\extracolsep{5pt}} ccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
named\_run\_type & is\_new & mean\_loss & boot\_CI\_low & boot\_CI\_high \\
\hline \\[-1.8ex]
EML & Trained & 0.014 & 0.011 & 0.017 \\
EML & Held out & 0.218 & 0.153 & 0.291 \\
Untrained EML network & Trained & 5.735 & 4.787 & 6.787 \\
Untrained EML network & Held out & 5.968 & 4.967 & 6.975 \\
\hline \\[-1.8ex]
\end{tabular}
\caption{Table for basic meta-learning, figure \ref{poly_basic_results}}
\end{table}
